Two applications of logic programming to Coq

### 2021-03-05 papers/types2020/revisions/cultures.tex
Updates to section 2.

Gentzen's discovered that the key to treating classical and
intuitionistic logics in the same proof system was the weakening and
contraction structural rules on the right-side of
sequents~\cite{gentzen35}.  This discovery lead him away from natural
deduction to multiple-conclusion sequent calculus, a very important
kind of proof system that has not been given a satisfactory
representation in type theory.  This same innovation of Gentzen also
opened the way to another key proof-theoretic discovery, that of
linear logic \cite{girard87tcs}.  Similarly, sequent calculus provides
an elegant presentation of linear logic while most treatments of
linear logic in type theory have difficulty treating linear negation
and the multiplicative disjunction  and its unit.

@@

Gentzen-style proofs are used to model computation in at least two
different ways.  The functional programming paradigm, following the
Curry-Howard correspondence, models computation abstractly as
$\beta$-reduction of natural deduction proof: that is, computation is
modeled as \emph{proof normalization}.  On the other hand, the logic
programming paradigm, following the notion of goal-directed proof
search, models computation abstractly as a regimented search for
cut-free sequent calculus proofs: that is, computation is modeled as
\emph{proof search}.

Proof search has features that are hard to capture in the proof
normalization setting.  In particular, Gentzen's \emph{eigenvariables}
are a kind of proof-level binder.  In the proof normalization setting,
such eigenvariables are instantiated during proof normalization.
However, during the search for cut-free proofs, eigenvariables do not
vary, and they are part of the syntax of terms and formulas.  As a
result, they can be used in the treatment of bindings in data
structures more generally.  Such eigenvariables can be used by the
logic programming paradigm to provide a natural, elegant, and powerful
computing approach to bindings within syntax.

>>>>

Gentzen's discovered that the key to treating classical and
intuitionistic logics in the same proof system was identifying the
structural rules of the weakening and contraction and
placing them on the right-side of sequents~\cite{gentzen35}.
This discovery lead him away from natural deduction to
multiple-conclusion sequent calculus, a very important kind of proof
system that has not been given a satisfactory representation in type
theory.  This same innovation of Gentzen's also opened the way to
another key proof-theoretic discovery, that of linear logic
\cite{girard87tcs}.  Similarly, sequent calculus provides an elegant
presentation of linear logic while most treatments of linear logic in
type theory have difficulty treating linear negation and the
multiplicative disjunction and its unit.

@@

Gentzen-style proofs are used to model computation in at least two
different ways.  The functional programming paradigm, following the
Curry-Howard correspondence, models computation abstractly as
$\beta$-reduction within natural deduction proofs: that is,
computation is modeled as \emph{proof normalization}.  On the other
hand, the logic programming paradigm, following the notion of
goal-directed proof search, models computation abstractly as a
regimented search for cut-free sequent calculus proofs: that is,
computation is modeled as \emph{proof search}.

Proof search has features that are hard to capture in the proof
normalization setting.  In particular, Gentzen's \emph{eigenvariables}
are a kind of proof-level binder.  In the proof normalization setting,
such eigenvariables are instantiated during proof normalization.
However, during the search for cut-free proofs, eigenvariables do not
vary, and they are part of the syntax of terms and formulas.  As a
result, they can be used in the treatment of bindings in data
structures more generally.  Such eigenvariables can be used by the
logic programming paradigm to provide a natural and powerful
approach to computing with bindings within syntax.
